Negative and Fractional Types

### 2012-06-27 rational_ext.tex
vector spaces

\newcommand{\fname}[1]{\ulcorner #1 \urcorner}
\newcommand{\fconame}[1]{\llcorner #1 \lrcorner}

@@

%%%%%%%%%%%%%%%%%
\subsection{The Powerset Monad}

In order to interpret fractional types, we first define a monad {{S}} over
the {{Pi}}-category of finite sets and bijections.

>>>>

\renewcommand{\vec}[2]{\begin{bmatrix} #1 \\ #2 \end{bmatrix}}
\newcommand{\Fpx}[1]{\mathbf{F}_{#1}}
\newcommand{\modulusp}[1]{\left | #1 \right |}
\newcommand{\fname}[1]{\ulcorner #1 \urcorner}
\newcommand{\fconame}[1]{\llcorner #1 \lrcorner}

@@

%%%%%%%%%%%%%%%%%
\subsection{Vector Spaces}

In order to interpret fractional types, we first lift the values and
combinators of {{Pi}} to the world of vector spaces. Intuitively:
\begin{itemize}
\item each type {{b}} is now interpreted as a vector space of dimension
  $\modulusp{b}$ over the field $\Fpx{2}$; the elements of such a vector
  space are column vectors
\item each value {{v_i : b}} is interpreted as column vector:
\[\begin{array}{r@{~}l}
\begin{array}{r} 
v_0 \\ v_1 \\ \vdots \\ v_i \\ \vdots \\ v_n 
\end{array} & 
\begin{bmatrix}
0 \\ 0 \\ \vdots \\ 1 \\ \vdots \\ 0
\end{bmatrix}
\end{array}\]
\item each combinator {{c : b1 <-> b2}} is interpreted as a matrix whose
  columns are indexed by the values of type {{b1}} and whose rows are indexed
  by the values of type {{b2}}. The matrix is \emph{orthogonal} with each row
  and column containing exactly one occurrence of the value 1.
\end{itemize}

Once we are in the world of vector spaces, the fractional type {{1/b}} is
interpreted as the \emph{dual} to the vector space denoted by {{b}}. The
elements of this dual vector space can be represented as \emph{row vectors}.

Multiplying a row vector by a column gives a scalar: this is eta and epsilon.

\newpage
